Kernel Square Loss Exemplar Machines

### 2016-11-13 sec_kernel_methods.tex
up to section 4.

Assuming our reproducing kernel is semidefinite positive, 
%\footnote{\textit{i.e.} for any $z$ in $\RR^n$, $z^TKz\ge 0$. This is equivalent to all eigenvalues of $K$ being non-negative.} 
$K$ is a semidefinite positive matrix and can be decomposed as $K=BB^T$.
%, with $B$ a rank $r$ upper triangular matrix of size $n\times r$. 
Using this factorization, the kernelized problem can be expressed as 
\begin{equation}
\min_{\beta\in\RR^r,\nu\in\RR} \left( \dfrac{1}{n}\sumi l(y_i, b_i^T\beta+\nu)+\dfrac{\lambda}{2}\|\beta\|^2\right), \label{beta:first}
\end{equation}
where $b_i^T$ denotes the $i$-th row of $B$ and $r$ is the number of columns of $B$. 
If $(\beta^\star, \nu^\star)$ is the solution Equation (\ref{beta:first}), the corresponding vector $\alpha^\star$ (or, more correctly, \textit{a} corresponding vector of dimension $n\geq r$) 
can be computed by $\alpha^\star=P\beta^\star$, where $P$ is the pseudoinverse of $B^T$.

@@

\begin{lemma} The augmented kernel matrix $K'$ can be factorized as $K'= B'B'^T$ with
%\begin{align}
%K'&= B'B'^T,\quad\text{where}\quad
%B'=\begin{bmatrix}
%u & v^T\\0 & B
%\end{bmatrix}\\
%v&=B^\dagger k_0,\,\, u=\sqrt{k_{00}-||v||^2}.\label{eq:lemma1}
%\end{align}
\begin{equation}
B'=\begin{bmatrix}
u & v^T\\0 & B
\end{bmatrix},~
v = B^\dagger k_0,~ u=\sqrt{k_{00}-||v||^2}.
\label{eq:lemma1}
\end{equation}
\end{lemma}\label{lemma1}
\begin{proof}
For $B'$ defined by (\ref{eq:lemma1}), we have that
\begin{equation}
B'B'^T = 
\begin{bmatrix} u^2+\|v\|^2 & v^TB^T\\ 
Bv& BB^T\end{bmatrix} 
=\begin{bmatrix} k_{00} & v^TB^T\\ 
Bv& K\end{bmatrix} .
\end{equation}
Since $K'$ is positive semidefinite, $k_0$ must lie in the column space $\mathcal{B}$ of $B$.
%\footnote{
Indeed, if we suppose $k_0$ does not belong to $\mathcal{B}$, then it can be decomposed uniquely as $k_0=s+t$, $s\in\mathcal{B}$ and  $t\in\mathcal{B}^\perp$, with $t\ne 0$. In one hand, $K'$ being semidefinite positive implies that $[1, -at^T]K'[1; -at]=k_{00}-2a\|t\|^2\ge 0$ for all real value $a$. In the other hand, for $a$ large enough, $k_{00}-a\|t\|^2\le 0$, which is a contradiction.
%} 
Hence $v=B^\dagger k_0$ is an exact solution of $Bv=k_0$. The fact that $k_{00}-\|v\|^2$ is non-negative comes from the fact that the Schur complement $K-k_0 k_0^T / k_{00}$ of $k_{00}$ in $K'$ is itself positive semidefinite.
%\footnote{
Indeed, since the matrix $k_{00}K-k_0k_0^T=B(k_{00}\mathrm{Id}_r-vv^T$ is also positive semidefinite. Thus $v^T(k_{00}\mathrm{Id}_r-vv^T)v = \|v\|^2(k_{00}-\|v\|^2)\ge 0$. \qedhere
%}
\end{proof}

This lemma allows us to add a positive sample to Eq. (\ref{beta:first}).
%$\beta^\star(x_0,X), \nu^\star(x_0,X) = \underset{\beta\in\RR^{r+1}, \nu\in\RR}{argmin}J'(\beta, \nu)$, for
With a positive exemplar, Eq. (\ref{beta:first}) now reads
\begin{align}
 \dfrac{1}{n}l(1, b_0'^T\beta+\nu) +\dfrac{1}{n}\sumi l(-1,b_i'^T\beta+\nu)
+\dfrac{\lambda}{2}\|\beta\|^2,\label{beta:final}
\end{align}
with $b_i'^T$ being the $(i+1)$-th row of $B'$, $i$ in $\{0,1,...,n\}$. In particular, $b_0'=[u; \ v]$ and, for $i>0$, $b_i'=[0; \ b_i]$. The solution $(\beta^\star,\nu^\star)$ in $\RR^{r+1}\times\RR$ can be computed just as before by Equation (\ref{omega:solution}), replacing $x_0$ by $b_0'$, $\mu$ by $\mu' = \frac{1}{n}\sum_{i=1}^n b_i'$ and $X$ by the $(r+1)\times n$ matrix $Q$ of columns $b_1', b_2',...,b_n'$. $\alpha^\star$ is now calculated as
$\alpha^\star=P'\beta^\star$, where $P' = [u^{-1} \ 0^T; -u^{-1}Pv \ P]$ is the pseudoinverse of $B'^T$. 
%This can be expressed by the linear system

@@

For a given image whose descriptor is $x_0$, we need to store $x_0$, $\hat{\beta}$ and $v$ to calculate its matching score to whichever other image for SLEM. Each image depends therefore on a vector of $p+2r$ dimensions.

>>>>

Assuming our reproducing kernel is semidefinite positive, 
%\footnote{\textit{i.e.} for any $z$ in $\RR^n$, $z^TKz\ge 0$. This is equivalent to all eigenvalues of $K$ being non-negative.} 
$K$ is a semidefinite positive matrix and can be decomposed as $K=BB^T$.
%, with $B$ a rank $r$ upper triangular matrix of size $n\times r$. 
Using this factorization, the kernelized problem can be expressed as 
\begin{equation}
\min_{\beta\in\RR^r,\nu\in\RR} \left( \dfrac{1}{n}\sumi l(y_i, b_i^T\beta+\nu)+\dfrac{\lambda}{2}\|\beta\|^2\right), \label{beta:first}
\end{equation}
where $b_i^T$ denotes the $i$-th row of $B$ and $r$ is the number of columns of $B$. 
If $(\beta^\star, \nu^\star)$ is the solution of (\ref{beta:first}), the corresponding vector $\alpha^\star$ (or, more correctly, \textit{a} corresponding vector of dimension $n\geq r$) 
can be computed by $\alpha^\star=P\beta^\star$, where $P$ is the pseudoinverse of $B^T$.

@@

\begin{lemma} The augmented kernel matrix $K'$ can be factorized as $K'= B'B'^T$ with
%\begin{align}
%K'&= B'B'^T,\quad\text{where}\quad
%B'=\begin{bmatrix}
%u & v^T\\0 & B
%\end{bmatrix}\\
%v&=B^\dagger k_0,\,\, u=\sqrt{k_{00}-||v||^2}.\label{eq:lemma1}
%\end{align}
\begin{equation}
B'=\begin{bmatrix}
u & v^T\\0 & B
\end{bmatrix},~
v = B^\dagger k_0,~ u=\sqrt{k_{00}-||v||^2},
\label{eq:lemma1}
\end{equation}
where $B^\dagger$ is the pseudoinverse of $B$.
\end{lemma}\label{lemma1}
\begin{proof}
For $B'$ defined by (\ref{eq:lemma1}), we have that
\begin{equation}
B'B'^T = 
\begin{bmatrix} u^2+\|v\|^2 & v^TB^T\\ 
Bv& BB^T\end{bmatrix} 
=\begin{bmatrix} k_{00} & v^TB^T\\ 
Bv& K\end{bmatrix} .
\end{equation}
Since $K'$ is positive semidefinite, $k_0$ must lie in the column space $\mathcal{B}$ of $B$.
%\footnote{
Indeed, if we suppose $k_0$ does not belong to $\mathcal{B}$, then it can be decomposed uniquely as $k_0=s+t$, $s\in\mathcal{B}$ and  $t\in\mathcal{B}^\perp$, with $t\ne 0$. In one hand, $K'$ being semidefinite positive implies that $[1, -at^T]K'[1; -at]=k_{00}-2a\|t\|^2\ge 0$ for all real value $a$. In the other hand, for $a$ large enough, $k_{00}-a\|t\|^2\le 0$, which is a contradiction.
%} 
Hence $v=B^\dagger k_0$ is an exact solution of $Bv=k_0$. The fact that $k_{00}-\|v\|^2$ is non-negative comes from the fact that the Schur complement $K-k_0 k_0^T / k_{00}$ of $k_{00}$ in $K'$ is itself positive semidefinite.
%\footnote{
Indeed, since the matrix $k_{00}K-k_0k_0^T=B(k_{00}\mathrm{Id}_r-vv^T)$ is also positive semidefinite. Thus $v^T(k_{00}\mathrm{Id}_r-vv^T)v = \|v\|^2(k_{00}-\|v\|^2)\ge 0$. \qedhere
%}
\end{proof}

This lemma allows us to add a positive sample to Eq. (\ref{beta:first}).
%$\beta^\star(x_0,X), \nu^\star(x_0,X) = \underset{\beta\in\RR^{r+1}, \nu\in\RR}{argmin}J'(\beta, \nu)$, for
With a positive exemplar, Eq. (\ref{beta:first}) now reads
\begin{align}
 \dfrac{1}{n}l(1, b_0'^T\beta+\nu) +\dfrac{1}{n}\sumi l(-1,b_i'^T\beta+\nu)
+\dfrac{\lambda}{2}\|\beta\|^2,\label{beta:final}
\end{align}
with $b_i'^T$ being the $(i+1)$-th row of $B'$, $i$ in $\{0,1,...,n\}$. In particular, $b_0'=[u; \ v]$ and, for $i>0$, $b_i'=[0; \ b_i]$. The solution $(\beta^\star,\nu^\star)$ in $\RR^{r+1}\times\RR$ can be computed just as before by Equation (\ref{omega:solution}), replacing $x_0$ by $b_0'$, $\mu$ by $\mu' = \frac{1}{n}\sum_{i=1}^n b_i'$ and $X$ by the $(r+1)\times n$ matrix $Q$ of columns $b_1', b_2',...,b_n'$. Solution $\alpha^\star$ is now calculated as
$\alpha^\star=P'\beta^\star$, where $P' = [u^{-1} \ 0^T; -u^{-1}Pv \ P]$ is the pseudoinverse of $B'^T$. 
%This can be expressed by the linear system

@@

For a given image whose descriptor is $x_0$, we need to store $x_0$, $\hat{\beta}$ and $v$ to calculate its matching score to whichever other image for SLEM. Each image depends therefore on a vector of $p+2r$ dimensions \PP{Remind what are $p$ and $r$, I lost the former!}.
